---
title: "Music Analysis"
author: "Yonatan Amaru"
date: "6/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Creating metrics
in order to create more fitting metrics we will analyze the music data by country happiness
1. reshaping the data:
```{r}
merged_df <- merge(top10mean, happ_scaled, by = 0)
colnames(merged_df)[1] <- "country"
merged_df <- arrange(merged_df, desc(`Ladder score`))
```

```{r}
merged_df.long <- melt(merged_df, id = "Ladder score", measure = c("energy_1", "valence_1", "danceability_1", "acousticness_1"))
ggplot(merged_df.long, aes(`Ladder score`, value, colour = variable)) + geom_point() + geom_smooth(method='lm', formula= y~x)
```

nada :(

```{r}
simple.reg <- lm(`Ladder score`~valence_1, data=merged_df)
summary(simple.reg)
```
the Multiple R-squared is very low, not a very good indicator

```{r}
plot(merged_df[c(3, 8,2,13, 10,12)])
```
```{r}
double.reg <- lm(`Healthy life expectancy`~valence_1+danceability_1+energy_1, data=merged_df)
summary(double.reg)
```

```{r}
corrplot(cor(merged_df[-1]), type = "lower", method = "ellipse", tl.cex = 0.9)
```
no strong correlation

let present some of the data mapped in order
artist diversity, dance variance

```{r}
div_top50 <- top50country %>%
  group_by(country) %>% 
  summarise(unique = n_distinct(artists), n = n(), .groups = 'drop', sd_dance=sd(danceability)) %>% 
  mutate(artist_diversity = unique/n)

div_top50["sd_dance"] <- (div_top50["sd_dance"]-min(div_top50["sd_dance"]))/(max(div_top50["sd_dance"])-min(div_top50["sd_dance"])) 
div_top50
```
range of valence
```{r}

merged_df.long <- melt(merged_df, id = "energy_1", measure = c("Ladder score", "Healthy life expectancy", "Freedom to make life choices"))
ggplot(merged_df.long, aes(energy_1, value, colour = variable)) + geom_point() + geom_smooth(method='lm', formula= y~x)

```
```{r}
merged_df_v2 <- merge(merged_df, div_top50[-c(2, 3)], by = 1)
```
```{r}
corrplot(cor(merged_df_v2[-1]), type = "lower", method = "ellipse", tl.cex = 0.9)
```
```{r}
merged_df.long <- melt(merged_df_v2, id = "Logged GDP per capita", measure = c("sd_dance", "artist_diversity"))
ggplot(merged_df.long, aes(`Logged GDP per capita`, value, colour = variable)) + geom_point() + geom_smooth(method='lm', formula= y~x)
```

```{r}
double.reg <- lm(`Logged GDP per capita`~artist_diversity+sd_dance, data=merged_df_v2)
summary(double.reg)
```

```{r}
colnames(merged_df_v2)
```
```{r}
# CLEANING MAIN GENRES
top50country %>% 
    select('country', 'genre') %>%
       mutate(genre = replace(genre, str_detect(genre, 'pop'), 'pop'),
              genre = replace(genre, str_detect(genre, 'latin'), 'latin'),
              genre = replace(genre, str_detect(genre, 'rock'), 'rock'),
              genre = replace(genre, str_detect(genre, 'blues'), 'blues'),
              genre = replace(genre, str_detect(genre, 'rap'), 'rap'),
              genre = replace(genre, str_detect(genre, 'edm'), 'edm'),
              genre = replace(genre, str_detect(genre, 'r&b'), 'r&b'),
              genre = replace(genre, str_detect(genre, 'country'), 'country'),
              genre = replace(genre, str_detect(genre, 'house'), 'house'),
              genre = replace(genre, str_detect(genre, 'indie'), 'indie'),
              genre = replace(genre, str_detect(genre, 'soul'), 'soul'),
              genre = replace(genre, str_detect(genre, 'reggae'), 'reggaeton'),
              genre = replace(genre, str_detect(genre, 'dance'), 'dance'),
              genre = replace(genre, str_detect(genre, 'hip hop'), 'hip hop')) -> top_genres
  
```

```{r}
genre_count <- top_genres %>%
  group_by(country, genre) %>%
  summarise(n = n()) %>%
  arrange(country, desc(n)) 
  

pop <- genre_count %>% 
  filter(genre == "pop") %>%
  mutate(pop = n/50)

hiphop <- genre_count %>% 
  filter(genre == "hip hop") %>%
  mutate(hip_hop = n/50)

```
```{r}
merged_df_v3 <- merge(merged_df_v2, pop[-c(2, 3)], by = 1)
merged_df_v3 <- merge(merged_df_v3, hiphop[-c(2, 3)], by = 1)
```

```{r}
merged_df.long <- melt(merged_df_v3, id = "Logged GDP per capita", measure = c("pop"))
ggplot(merged_df.long, aes(`Logged GDP per capita`, value, colour = variable)) + geom_point() + geom_smooth(method='lm', formula= y~x)
```
```{r}
merged_df.long <- melt(merged_df_v3, id = "Healthy life expectancy", measure = c("hip_hop"))
ggplot(merged_df.long, aes(`Healthy life expectancy`, value, colour = variable)) + geom_point() + geom_smooth(method='lm', formula= y~x)
```

maybe not a linear model? -x^2?


```{r}
simple.reg <- lm(`Logged GDP per capita`~pop, data=merged_df_v3)
summary(simple.reg)
```

2. Scraping lyrics off azlyrics.biz

```{r, lyric_func, eval=FALSE}

lyric_scraper <- function(url) {
  m <- read_lines(url)
  giveaway <- '<div id="konten"></div>'
  start <- grep(giveaway, m) + 1
  end <- grep('<div', m[start[1]:length(m)])[1] + start[1]
  lyrics <- paste(gsub("<br />|</div>|</p>", "", m[start[1]:end]), collapse = " ")
  return(lyrics)
  
  #TODO: fix end
}

#TODO add multiple artists
name_builder <- function(name, artist) {
  # clean name
  song_clean  <- tolower(gsub("[[:punct:]]","",name))
  song_clean  <- gsub("\\s","-",song_clean)
  # clean artist
  artist_clean  <- tolower(gsub("[[:punct:]]","",artist))
  artist_clean  <- gsub("\\s","-",artist_clean)
  name <- paste(artist_clean, song_clean, sep="-")
  sep_url <- c("https://azlyrics.biz/", substring(artist_clean, 1, 1), "/", artist_clean, "-lyrics/", name, "-lyrics/")
  url <- paste(sep_url, collapse = "")
  return(url)
}
```

```{r}
lyric_scraper(name_builder("Running Up That Hill (A Deal With God)", "kate bush"))
```
```{r}

lyric_scraper <- function(url) {
    out <- tryCatch(
        {
            m <- read_lines(url)
            giveaway <- '<div id="konten"></div>'
            start <- grep(giveaway, m) + 1
            end <- grep('<div', m[start[1]:length(m)[1]])[1] + start[1]
            lyrics <- paste(gsub("<br />|</div>|</p>", "", m[start[1]:end]), collapse = " ")
            # The return value of `readLines()` is the actual value 
            # that will be returned in case there is no condition 
            # (e.g. warning or error). 
            # You don't need to state the return value via `return()` as code 
            # in the "try" part is not wrapped inside a function (unlike that
            # for the condition handlers for warnings and error below)
        },
        error=function(cond) {
            # Choose a return value in case of error
            message(paste("URL does not seem to exist:", url))
            message("Here's the original error message:")
            message(cond)
            return(NA)
        },
        warning=function(cond) {

            # Choose a return value in case of warning
            message(paste("URL caused a warning:", url))
            message("Here's the original warning message:")
            message(cond)
            return(NULL)
        },
        finally={
        # NOTE:
        # Here goes everything that should be executed at the end,
        # regardless of success or error.
        # If you want more than one expression to be executed, then you 
        # need to wrap them in curly brackets ({...}); otherwise you could
        # just have written 'finally=<expression>' 
            message(paste("Processed URL:", url))
        }
    )    
    return(out)
}
```


next we will run through all here songs and scrape them of azlyrics:
```{r, azlyrics, eval=FALSE}
song <- c(top50country['name'])
artist <- c(top50country['artists'])

urls <- lapply(song, name_builder, artist)
```
```{r}
urls
```



We will now save it to the data file.
```{r, savelyrics, eval=FALSE}
adele_data$lyrics <- datalist
save(adele_data, file = "../data/adele_data.Rdata")
```

diversity - average distance from songs to what?
dominant genre to gdp
locality - hard to do?

testing OECD tests to test for correlation - Nitzan N Eden

```{r}
# OECD <- read.csv("../data/OECD.csv",  encoding = "UTF-8")
# OECD %>%
#   mutate(math.sd = str_replace_all(math.sd, "\\*|\\(|\\)", "")) %>%
#   mutate(reading.sd = str_replace_all(reading.sd, "\\*|\\(|\\)", "")) %>%
#   mutate(science.sd = str_replace_all(science.sd, "\\*|\\(|\\)", "")) %>%
#   mutate(math.sd = as.numeric(math.sd)) %>%
#   mutate(reading.score = as.numeric(reading.score)) %>%
#   mutate(reading.sd = as.numeric(reading.sd)) %>%
#   mutate(science.sd = as.numeric(science.sd)) -> OECD
# 
# write.csv(x=OECD, file="../data/OECD_clean")
OECD <- read.csv("../data/OECD_clean.csv",  encoding = "UTF-8")
```

```{r}
merged_df <- merge(rmse.score, OECD, by="country")
merged_df$rmse.score <- merged_df$rmse.score * 100

merged_df %>%
  mutate(reading.score = ifelse(is.na(reading.score),
                            median(reading.score, na.rm = TRUE), 
                            reading.score)) %>%
  mutate(reading.sd = ifelse(is.na(reading.sd),
                            mean(reading.sd, na.rm = TRUE), 
                            reading.sd)) -> merged_df
  
  
```

